% Introduction

This document presents a collection of mathematical problems known as \emph{coffin problems}:
they are challenging problems that in the 1960s, 1970s, and 1980s
were given at some Soviet universities' admission exams
with the goal of preventing the enrolment of students who were considered `undesirable',
category which mainly included people of Jewish origin.
Such students were administered these problems at their oral examinations.

\ifdefined\ONLYSOLUTIONS
  \ifdefined\NOSOLUTIONS
    % Empty document
  \else
    This is the solution document: each problem is immediately followed by its solution.
  \fi
\else
  \ifdefined\NOSOLUTIONS
    The solutions to the problems are not provided in this document.
  \else
    Solutions to the problems are provided in the second part of the document.
  \fi
\fi

\ifdefined\GENERALISED
  \ifdefined\ORIGINAL
    In the second half of the document, interesting generalisations of some of the problems are presented.
  \else
    The problems listed in this document are generalisations of the original problems.
  \fi
\fi

\ifdefined\UNCATEGORISED
\else
  The problems are categorised based on topic:
  each section lists problems that fall under a particular category.
\fi

The references from which the problems were sourced are listed at the end of the document.%
\ifdefined\NOSOURCES%
\else%
  Moreover, at each problem statement, the specific sources in which that problem is documented are cited.%
\fi%
\ifdefined\MOREREFS%
  Additional resources for further reading on the topic are also provided, after the main references.
\fi%
\ifdefined\FORUMREFS%
  References to user discussions are mentioned at the end of the document.
\fi
